Model of lens candidates from Space Warps CFHTLS

### 2016-07-04 paper2.tex
editing text

\newcommand{\figref}[1]{Figure \ref{fig:#1}}
\newcommand{\Figref}[1]{Figure \ref{fig:#1}}

@@

\begin{abstract}
This is a simple template for authors to write new MNRAS papers.
The abstract should briefly describe the aims, methods, and main results of the paper.
It should be a single paragraph not more than 250 words (200 words for Letters).
No references should appear in the abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

Space Warps \citep{2016MNRAS.455.1171M} found lens candidates
\citep{2016MNRAS.455.1191M}.

Modelling method \citep{2015MNRAS.447.2170K}.

\section{Methods}

\subsection{Stellar mass}

\subsection{Parameterisation of pixel models} \label{sec:parameter}
In order to fit the set of pixelated models to a single parameterised model, a program was written that took a parameterised function and subtracted from it the mean and the principle components of the data, which were calculated using classical Principle Component Analysis.
This created the residuals function.
The number of components defined as principle was varied to test how this affected the output, and it was found that using 5 principle components tended to give a reasonable approximation.
A masking function was added which selected only the data points that fell inside the image of the lens, and the principal components were clipped in order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 as, assuming that the data follows a Gaussian error distribution, almost all the values for the variance should lie between 2 and 3 standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the parameterised function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.

The parameterised model function was obtained from the gravitational potential of an isothermal ellipsoid mass distribution \cite{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of the Einstein ring, the ellipticity of the model and the angle of the ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these parameters were already known, it was possible to gain an estimate of the projected accuracy of the results, before applying the model to the candidate lensing galaxies.

@@

\subsection{HiRes / Subsampling of central region}
In a previous paper \cite{2015MNRAS.447.2170K} we identified the problem of the tendency of the models to be too shallow.
We pinned down the problem to the central region and are confident to have it fixed with enabling sub pixel sampling in the central area.
The software allows for subsampling parameters of radius of pixels $r_\text{subs}$ and amount of subpixels per pixel $n_\text{subs}$.
\Figref{subsampling} shows the results of different settings for $r_\text{subs}$ and $n_\text{subs}$.
We can see that even the computationally least expensive settings of $r_\text{subs}=1$ and $n_\text{subs}=3$ leads to drastically improved profiles.

@@

@@

\begin{figure}
  \inclfign{SW28}{ASW0007xrs_JHC3J2HYV7}
  \caption{SW28.  See text in Section~\ref{sec:examples} for details.}
  \label{fig:SW28}
\end{figure}

\begin{figure}
  \inclfign{SW05}{ASW0007k4r_AJIBCHQ6EM}
  \caption{SW05}
  \label{fig:SW05}
\end{figure}

@@

Each of Figures~\ref{fig:SW28}--\ref{fig:SW57} show six panels, as
follows.

@@

\item Lower left is synthetic image.
  Figures~\ref{fig:SW28}--\ref{fig:SW29} include source fitting.
  Figures~\ref{fig:SW42}--\ref{fig:SW57} uses a simple conical
  profile.

@@

\item Lower right shows stellar vs stellar mass.\todo{which population
  model?}

@@

\item Figures~\ref{fig:SW28}--\ref{fig:SW29} show the five most
  convincing lens candidates. SW28 shown in Figure~\ref{fig:SW28} is
  modelled as an arc with three blended images plus a saddle-point
  closer to the galaxy.  SW05 shown in Figure~\ref{fig:SW05} has four
  distinct images (two close to blending).  SW02, SW09 and SW29
  (Figures~\ref{fig:SW02}--\ref{fig:SW29}) have similar
  configurations, with a minimum furthest from the galaxy and three
  images blended into an arc.  All five would be massive ellipticals;
  SW05 is the most massive of all the candidates, with a galaxy-group
  scale mass.

\item Figure~\ref{fig:SW42} shows a candidate that modelling indicates
  is doubtful.  The images appears to have a similar morphology to the
  last three mentioned.  The implied mass, however, is implausibly
  high.\todo{Check photometry of this candidate.}

\item Figures~\ref{fig:SW58} and \ref{fig:SW19} appear plausible lens
  candidates.  Their morphology is similar to SW28, but the
  saddle-point counterimage is not visible.  We consider these lenses
  plausible but less convincing.

\item Figures~\ref{fig:SW36} and \ref{fig:SW57} shows {\em unclear\/}
  cases where modelling failed.

@@

@@

The canidates that looked the most promising and left little doubt about being lenses were grouped into the category ``convincing''.
In Figures \ref{fig:SW02} to \ref{fig:SW29} we show all the convincing candidates.
Following we list the parameters, obtained applying the technique described in section~\ref{sec:parameter}

@@

\section{Discussions}

>>>>

\newcommand{\figref}[1]{Figure \ref{fig:#1}}

@@

\begin{abstract}
We report modelling follow-up of recently-discovered
gravitational-lens candidates in the CFHT Legacy Survey.  Lens
modelling was done by a small group of specially-interested volunteers
from the Space~Warps citizen-science community who originally found
the candidate lenses.  Models are categorised according to seven
qualitative and quantitive points.  Also included are some
improvements to the modelling software used (SpaghettiLens) are
presented, and discussion of strategies for scaling to future surveys
with more and frequent discoveries.

The candidates successfully modelled are all galaxies, with inferred
lensing masses ranging from $\sim10^{11}M_\odot$ to $>10^{13}M_\odot$.
Stellar masses have also been estimated, using photometry from the
CFHTLS pipeline and stellar-population models.  The trend well-known
in nearby galaxies, that star-formation efficiency is maximal for
total masses of $\approx10^{12}M_\odot$ and reduces for both lower and
higher masses, is discernable also in these much more distant
galaxies.
\end{abstract}

\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

A passing remark in the {\em Feynman Lectures}\footnote{
  \cite{1963flp..book.....F}} --- ``if you are looking for a good
problem, the exact details of how the arms are formed and what
determine the shapes of these galaxies has not been worked out'' ---
tells us that circa 1960, galaxy dynamics was expected be an easy
research puzzle, whose solution was just around the corner.  Yet a
half-century later, galaxies are not nearly as well-understood as
stars were then.  Two complications, which are clear now but were not
so clear then, are: (i)~dynamical chaos, because galaxies are not
spherical, and (ii)~dark matter.

The CFHTLS \citep{2012SPIE.8448E..0MC} has been a rich source of
lenses by arc-finders, by machine learning \citep{2016arXiv160504309P}
and by visual inspection \citep{2016MNRAS.455.1171M}.  The latter is
of particular interest here.

\section{Lens Modelling}

@@

\subsection{HiRes / Subsampling of central region}
In a previous paper \cite{2015MNRAS.447.2170K} we identified the problem of the tendency of the models to be too shallow.
We pinned down the problem to the central region and are confident to have it fixed with enabling sub pixel sampling in the central area.
The software allows for subsampling parameters of radius of pixels $r_\text{subs}$ and amount of subpixels per pixel $n_\text{subs}$.
\figref{subsampling} shows the results of different settings for $r_\text{subs}$ and $n_\text{subs}$.
We can see that even the computationally least expensive settings of $r_\text{subs}=1$ and $n_\text{subs}=3$ leads to drastically improved profiles.

@@

\subsection{Parameterisation of pixel models} \label{sec:parameter}
In order to fit the set of pixelated models to a single parameterised model, a program was written that took a parameterised function and subtracted from it the mean and the principle components of the data, which were calculated using classical Principle Component Analysis.
This created the residuals function.
The number of components defined as principle was varied to test how this affected the output, and it was found that using 5 principle components tended to give a reasonable approximation.
A masking function was added which selected only the data points that fell inside the image of the lens, and the principal components were clipped in order to keep the values inside the region of the ensemble of models.
Any value higher than the clip was set to be the clip value.
This was chosen to be 2.5 as, assuming that the data follows a Gaussian error distribution, almost all the values for the variance should lie between 2 and 3 standard deviations from the mean.
Minimising the residuals function produces the set of parameters that fit the parameterised function to the original pixelated ensemble most closely.
A least squares fit was used to perform this minimisation.
The parameterised model function was obtained from the gravitational potential of an isothermal ellipsoid mass distribution \cite{2001astro.ph..2341K}.
This model is frequently used to describe gravitational lenses as it tends to fit well with observations.
The isothermal ellipsoid model outputs three useful parameters: the radius of the Einstein ring, the ellipticity of the model and the angle of the ellipticity from the vertical, giving the orientation of the galaxy.
By applying this model to simulated lenses for which the values of these parameters were already known, it was possible to gain an estimate of the projected accuracy of the results, before applying the model to the candidate lensing galaxies.

\section{Stellar mass}

@@

\begin{figure}
  \inclfign{SW05}{ASW0007k4r_AJIBCHQ6EM}
  \caption{SW05.  See text in Section~\ref{sec:examples} for details.}
  \label{fig:SW05}
\end{figure}

\begin{figure}
  \inclfign{SW28}{ASW0007xrs_JHC3J2HYV7}
  \caption{SW28}
  \label{fig:SW28}
\end{figure}

@@

Each of Figures~\figref{SW05}--\figref{SW57} show six panels, as
follows.

@@

\item Lower left is synthetic image.
  Figures~\figref{SW05}--\figref{SW29} include source fitting.
  Figures~\figref{SW42}--\figref{SW57} uses a simple conical
  profile.

@@

\item Lower right shows stellar vs stellar mass.\todo{which population
  model?}  Abundance matching \citep{2010ApJ...710..903M}.

@@

\item Figures~\figref{SW28}--\figref{SW29} show the five most
  convincing lens candidates. SW28 shown in Figure~\figref{SW28} is
  modelled as an arc with three blended images plus a saddle-point
  closer to the galaxy.  SW05 shown in Figure~\figref{SW05} has four
  distinct images (two close to blending).  SW02, SW09 and SW29
  (Figures~\figref{SW02}--\figref{SW29}) have similar
  configurations, with a minimum furthest from the galaxy and three
  images blended into an arc.  All five would be massive ellipticals;
  SW05 is the most massive of all the candidates, with a galaxy-group
  scale mass.

\item Figure~\figref{SW42} shows a candidate that modelling indicates
  is doubtful.  The images appears to have a similar morphology to the
  last three mentioned.  The implied mass, however, is implausibly
  high.\todo{Check photometry of this candidate.}

\item Figures~\figref{SW58} and \figref{SW19} appear plausible lens
  candidates.  Their morphology is similar to SW28, but the
  saddle-point counterimage is not visible.  We consider these lenses
  plausible but less convincing.

\item Figures~\figref{SW36} and \figref{SW57} shows {\em unclear\/}
  cases where modelling failed.

@@

Table~\ref{tab:models}

\begin{enumerate}
\item Image morphology.
\item Whether images unblended.
\end{enumerate}

@@

The candidates that looked the most promising and left little doubt about being lenses were grouped into the category ``convincing''.
In Figures \figref{SW02} to \figref{SW29} we show all the convincing candidates.
Following we list the parameters, obtained applying the technique described in section~\ref{sec:parameter}

@@

\section{Discussion}
